Working with the Poisson distribution

### 2021-08-30 text/21/poisson-inferences.Rmd
Updated poisson-inferences.md

where T is known, but $lambda$ is unknown, then a 95% confidence interval for $lambda is

@@

@@

X ~ Pois(T$\lambda$) and

Y=ln(X), then

E[Y] $\approx$ ln($\lambda T$) and

V[Y] $\approx \frac{1}{\lambda T}$

@@

ln(X) $\pm 2 \sqrt{\frac{1}{X}}$

@@

@@

>>>>

where T is known, but $\lambda$ is unknown, then a 95% confidence interval for $\lambda$ is

@@

The general rule is that a nonlinear transformation leads to a similar change in the expected value, but a change in the variance that is related to the Taylor series expansion.

If X has an expected value, $\mu$ and a variance, $\sigma^2_x$, then g(X) can be written as

$g(X) \approx g(\mu)+(X-\mu)g'(\mu)$

which can be used to show that

$Var(g(X))=(g'(\mu))^2 \sigma^2_x$

@@

X ~ Pois($\lambda T$) and

$Y = g(X) = ln(X)$, then

$g'(X)=\frac{1}{X}$, and

V[Y] $\approx (g'(\mu))^2 \sigma^2_x = (\frac{1}{\lambda T})^2 \lambda T = \frac{1}{\lambda T}$

@@

$ln(X) \pm 2 \sqrt{\frac{1}{X}}$

@@

```{r echo=FALSE}
x <- 17
n <- 1989
lo <- x-2*sqrt(x)
hi <- x+2*sqrt(x)
log_lo <- log(x)-2*sqrt(1/x)
log_hi <- log(x)+2*sqrt(1/x)
exp_lo <- exp(log_lo)
exp_hi <- exp(log_hi)
```

The Pennsylvania Department of Health has a web page oulining calculations for rates with two simple examples. The infant mortality rate in Butler County in 1984 was 17 deaths out of 1,989 births. This is actually a proportion, but let's treat it as a rate here.

The death rate is 17/ 1989 =`r 17/1989`. To avoid dealing with small numbers, let's quote this as a rate per 1,000 live births. 1000 * 17/ 1989 = `r 1000*17/1989`. if we wanted a confidence interval for $\lambda T$, it would be

17 $\pm$  2 $\sqrt{17}$ = `r lo` to `r hi`

Divide everything by 1,989 and multiply by 1,000 to get

`r lo/1.989` to `r hi/1.989` which you should round to get 
`r round(lo/1.989, 1)` to `r round(hi/1.989, 1)`

Let's rework this using the better formula.

The confidence limits on the log scale are

$ln(17) \pm 2 \sqrt{\frac{1}{17}}$ = `r log_lo` to `r log_hi`

Exponentiate these values to get

`r exp_lo` to `r exp_hi`

Divide by 1,989 and multiply by 1,000 to get

`r exp_lo/1.989` to `r exp_hi/1.989`

and don't forget to round this: `r round(exp_lo/1.989, 1)` to `r round(exp_hi/1.989, 1)`.

### References

@@

https://en.wikipedia.org/wiki/Taylor_expansions_for_the_moments_of_functions_of_random_variables

https://en.wikipedia.org/wiki/Delta_method

https://en.wikipedia.org/wiki/Poisson_distribution

https://en.wikipedia.org/wiki/Binomial_distribution#Poisson_approximation

http://www.pmean.com/definitions/or.htm

http://www.pmean.com/01/oddsratio.html

http://www.pmean.com/04/Rates.html
